---
title: "states_through_time_5000"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(ape)
library(phytools)
library(tidyverse)
library(stringr)
```
Import files

```{r}
data_path <- "./data/"
simmap_files <- list.files(data_path, pattern = "^simmap")
simmap_files_names <- str_replace(simmap_files, "\\.RData", "")
```

Summarizing the results of simmap. Save it because it takes some time to calculate.
Especially in case of high number of simulations. 

```{r}
for(i in 1:length(simmap_files)){
  cat(i, "\n")
  env <- new.env()
  nm <- load(paste(data_path, simmap_files[i], sep = ""), env)[1]
  simmap <- env[[nm]]
  XX <- describe.simmap(simmap, plot = FALSE)
  save(XX, file = paste(data_path, simmap_files_names[i], ".describe.RData", sep = ""))
  rm(list = c("simmap", "XX"))
}

```

#branching times

Creating a list with the branching times of the tree

```{r}
node_ages <- list()
for(i in 1:length(simmap_files)){
  cat(i, "\n")
  env <- new.env()
  nm <- load(paste(data_path, simmap_files[i], sep = ""), env)[1]
  simmap <- env[[nm]]
  node_ages[[i]] <- branching.times(simmap[[1]])
  rm(list = c("simmap"))
}
names(node_ages) <- simmap_files_names
```


Extract the "ace" object from the summarised simmap files

```{r}
describe_files <- list.files(data_path, pattern = "describe.RData")
ace_list <- list()
for(i in 1:length(describe_files)){
  cat(i, "\n")
  env <- new.env()
  nm <- load(paste(data_path, describe_files[i], sep = ""), env)[1]
  describe <- env[[nm]]
  ace_list[[i]] <- describe$ace
  rm(list = c("describe"))
}
names(ace_list) <- describe_files
```

```{r}
# Are the two list is in the same order?
names(ace_list)
names(node_ages)
# Yes, so go ahead and create a data.frame from the lists
ace_list <- lapply(ace_list, data.frame)
for(i in 1:length(ace_list)){
  ace_list[[i]]$nodes <- rownames(ace_list[[i]])
}
for(i in 1:length(ace_list)){
  ace_list[[i]]$node_ages <- node_ages[[i]][match(ace_list[[i]]$nodes, names(node_ages[[i]]))]
}
ace_all <- bind_rows(ace_list)
```

The present character state is needed as well

```{r}
tree_files <- list.files(data_path, pattern = "kronogram")
trees <- lapply(tree_files, function(x) read.tree(paste(data_path, x, sep = "")))
fb_types_df <- read_delim(paste0(data_path, "FB_types_corrected_df.txt"), delim = "\t", col_names = TRUE)
#checking the species names, some correction is needed in the example files...
mis_names_trees <- which(trees[[1]]$tip.label %in% fb_types_df$Species == FALSE)
mis_names_trait <- which(fb_types_df$Species %in% trees[[1]]$tip.label == FALSE)
trees[[1]]$tip.label[mis_names_trees]
fb_types_df$Species[mis_names_trait]
# The misspelled names are the same order so easy to change it
fb_types_df[mis_names_trait, "Species"] <- trees[[1]]$tip.label[mis_names_trees]
# Some modifications on character states' name are needed as well
fb_types_df2 <- fb_types_df
fb_types_df2[is.na(fb_types_df2$corrected), "corrected"] <- "Resupinate"
fb_types_df2[fb_types_df2$corrected %in% c("Pileate-stipitate", "Pileate-sessile"), "corrected"] <- "Cap_Fungi"
fb_types_df2[fb_types_df2$corrected %in% c("Clavaroid", "Coralloid"), "corrected"] <- "Coral_Clavar"
unique(fb_types_df2$corrected)
# Now the names are ok, lets modify the dataframe a bit more
fb_types_df2$nodes <- as.character(1:nrow(fb_types_df2))
fb_types_df2$node_ages <- rep(0, nrow(fb_types_df2))
fb_types_df2$prob <- rep(1, nrow(fb_types_df2))
fb_types_df3 <- fb_types_df2[,-c(1:4)]
fb_types_df4 <- spread(data = fb_types_df3, key = corrected, value = prob)
fb_types_df4[,3:ncol(fb_types_df4)][is.na(fb_types_df4[,3:ncol(fb_types_df4)])] <- 0
#colnames(fb_types_df4) <- c("nodes", "node_ages", "Clavaroid", "Coralloid", "Cyphelloid", "Gasteroid", "no_fb", "Pileate.sessile", "Pileate.stipitate", "Resupinate")

```

So the important thing up is that have a dataframe like fb_types_df4!

Next the dataframe have to be multiplied as many times as many trees you analyzed

```{r}
fb_types_df5 <- fb_types_df4
i  <- 1
while(i != length(trees)){
  fb_types_df5 <- bind_rows(fb_types_df5, fb_types_df4)
  i <- i + 1
}
#The following should give the number of trees you analyzed:
nrow(fb_types_df5) / length(trees[[1]]$tip.label)
```

Lets combine the ASR and the present states data, then cut to 100 sections of the timeline.

```{r}
ace_all2 <- bind_rows(ace_all, fb_types_df5)
min_ages <- min(ace_all2$node_ages)
max_ages <- max(ace_all2$node_ages) + 1
section_size <- (max_ages - min_ages) / 100
sections <- seq(from = min_ages, to = max_ages, by = section_size)
names(sections) <- as.character(1:101)
```


Asign a section (out of 100) to each of the ages.

```{r}
ace_all2$section <- rep(NA, nrow(ace_all2))
for(i in 1:nrow(ace_all2)){
  for(j in 1:100){
      if(ace_all2[i, "node_ages"] >= sections[j] & ace_all2[i, "node_ages"] < sections[j+1]){
      ace_all2[i, "section"] <- names(sections)[j]
    }  
  }
  if(i %% 1000==0) {
      # Print on the screen some message
      cat(paste0("iteration: ", i, "\n"))
   }
}
```



```{r}
ace4 <- group_by(ace_all2, section) %>% 
  summarise_if(is.numeric, funs(sum, n()))
ace5 <- ace4[,2:8] / ace4[,9:15]
ace5$sections <- ace4$section
ace6 <- gather(ace5, key = "states", value = "prob_ratio", 1:6)
ace6 <- ace6[2:ncol(ace6)]
ace6$sections <- as.numeric(ace6$sections)
```

```{r}
save(ace6, sections,  file = "state_through_time_finalfigure.RData")
```


```{r}
states <- unique(ace6$states)
states <- states[c(5,1,2:4,6)] # you can change the order of the data so the 
ace6$states2 <- factor(ace6$states, levels = states)
ggplot(ace6, aes(x = sections,y = prob_ratio)) + 
  geom_area(aes(fill=states2, group = states2, colour = colors), position='stack', linetype = "solid", 
            size = 1, colour = "black", alpha = 0.6) +
  scale_x_reverse(labels = c("0", as.character(round(sections[c(0, 25, 50, 75, 101)],digits = 0))), breaks = c(0, 25, 50, 75, 100)) +
  scale_fill_manual(values = c("#0000EE", "#CD2626", "#00B2EE", "#228B22", "#FFFF00", "#FFB90F"))+
  theme_classic()
```


